---
title: "LiveSession05"
author: "Jeremy Otsap"
date: "June 2, 2019"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

##loading Libraries
library(dplyr)
library(tidyverse)

```

## Question 1

2016 Child names

```{r echo=TRUE}
# import yob2016 file
y2016 <- read.table("data/yob2016.txt", header = F, sep = ";", col.names = c("FirstName","Gender","Occurances"))

# dimensions & structure of dataframe
str(y2016)
dim(y2016)
#preview
y2016[1:10,]

```


Finding the problem entry with "yyy" at the end

```{r echo=TRUE}

# Locating problem entry with 'yyy'
y2016[grep('yyy', y2016$FirstName),]

```


Remove duplicate row from data set

```{r echo=TRUE}

slice(y2016, -212) -> y2016
y2016[205:215,]

#sorting data frame by FirstName
y2016[order(y2016$FirstName),] -> y2016


```



## Question 2

For the last 10 child names of 2015 they all are male and all have a total of 5 occurances. Also looks like its already in alphabetical order?

```{r echo=TRUE}

# import yob2015 file
y2015 <- read.table("data/yob2015.txt", header = F, sep = ",", col.names = c("FirstName","Gender","Occurances"))

#last 10 rows
tail(y2015, 10)

#sorting 2015 data frame
y2015[order(y2015$FirstName),] -> y2015

```


Merging 2015 and 2016 childname data frames

```{r echo=TRUE}
# using merge to emulate outer join
# only names picked in 2015 AND 2016 will be in final dataset
merge(y2015, y2016, by = c("FirstName","Gender")) -> final

colnames(final) <- c("Name","Gender","Occurances2015","Occurances2016")

```


## Question 3

Data summary of merged data sets showing most top 10 most popular names across 2015 and 2016

```{r echo=TRUE}

# create Total column
final$Total <- final$Occurances2015 + final$Occurances2016

# sorting by largest Total
final[order(final$Total, decreasing = T),] -> finaltotal

head(finaltotal, 10)

```

Top 10 girl names

```{r echo=TRUE}

# all girl names
finaltotal[finaltotal$Gender == 'F',] -> finalgirl
head(finalgirl[,c(1,5)],10)

# Export to CSV
finalgirl[,c(1,5)] -> finalgirlfile
write.csv(finalgirlfile, file = "data/finalgirl.csv", col.names = T)


```


## Question 4

####GitHub Address  
####https://github.com/jotsap/DDS6306_Unit05












